---
title: "CLARA divison analysis (II): inter SMU and profiles"
author: "Alberto Lázaro-López"
date: "3/03/2020"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
abstract: Just for covariates collection 2
bibliography: ../ref/dicsm.bib
---

```{r core, include=FALSE }
for (.i in paste0("src/", c("spatial", "dbms", "core", "raster", "nbclust_par"), ".R")) {
  source(.i)
}

proj <- .proj_envi("division") %>% 
  path_file() %>% 
  .proj_subdir()
```


# Introducción

Just for division from CLARA by covariates collection 2.

```{r, eval=TRUE}
div <- list("met" = "clara", 
            "sel" = 2)

proj[["data"]] <- path("data", paste0("division-", paste(div$met, div$sel, sep = "_")))
``` 

```{r}
smu1_set <- RPostgres::dbGetQuery(conGIS, glue_sql("
  SELECT DISTINCT smu1_id
  FROM division.{`paste(div$met, div$sel, 'div', sep = '_')`}", 
  .con = conGIS)
  )

smu1_set <- smu1_set %>% 
  mutate(smu1_id = tolower(smu1_id),
         class = path(proj$res, paste("clara", div$sel, sep = "_"), paste("clara", div$sel, smu1_id, sep = "-"), ext = "rds")) %>% 
  arrange(smu1_id)
```


# Extracción de centroides

Representación de los grupos a través de sus centroides.

```{r}
centroids <- foreach(
  i = smu1_set$smu1_id, 
  .final = function(l) setNames(l, smu1_set$smu1_id),
  .errorhandling = "pass") %dopar% {
    
    smu1_set %>% 
      filter(smu1_id == i) %>% 
      select(class) %>% 
      flatten_chr() %>% 
      readRDS() %>%
      .$medoids %>% 
      as_tibble() %>%
      mutate(grp = row_number()) %>% 
      select(grp, everything())
}

centroids <- centroids %>% 
  bind_rows(.id = "smu1_id") %>% 
  mutate_if(is.numeric, round, digits = 3) %>% 
  mutate(ctd_id = paste(smu1_id, grp, sep = "")) %>% 
  select(ctd_id, everything())

ctd_mat <- centroids %>% 
  magrittr::set_rownames(.$ctd_id) %>% 
  select(-smu1_id, -grp, -ctd_id) %>% 
  as.matrix()
```


## Hierarchical clustering

Diagrama - árbol bottom up

* Distance: Euclidean, Mahalanobis, Correlation

```{r}
dist_set <- c("Euclidean", "Mahalanobis", "Correlation")

ctd_dist <- foreach(i = dist_set, 
                    .final = function(l) setNames(l, dist_set)) %do% {
  proxy::dist(ctd_mat, method = i)
}
```

* Linkage: ward.D2 ~ average ~ complete

```{r}
ctd_hclust <- foreach(i = dist_set, 
                      .final = function(l) setNames(l, dist_set)) %do% {
  hclust(ctd_dist[[i]], method = "ward.D2")  
}
```

* Comparing dendogram

```{r}
ctd_dendro <- foreach(i = dist_set, 
                      .final = function(l) setNames(l, dist_set)) %do% 
{
  ctd_hclust[[i]] %>%
    as.dendrogram() 
    #dendextend::as.ggdend(offset_labels = 2) %>% 
    #ggplot()
}

#ggsave(path("gis", "test", ext = "pdf"), ctd_dendro$Euclidean, 
#       width = 15)

#' Run directly in the console
for (i in names(ctd_dendro)) {
  pdf(path(proj$img, paste("dendro", i, sep = "-"), ext = "pdf"), 
      width = 12, 
      height = 35)
  plot(ctd_dendro[[i]], horiz = TRUE)
  dev.off()
}
```


```{r}
dendro_comb <- arrangements::combinations(dist_set, k = 2, replace = FALSE)
dendro_names <- apply(dendro_comb, 1, function(r) paste(r, collapse = "_"))
ctd_tangle <- foreach(i = dendro_comb[, 1], 
                      j = dendro_comb[, 2], 
                      .final = function(l) setNames(l, dendro_names)) %do% {
  
  tangleidx <- round(dendextend::entanglement(ctd_dendro[[i]], ctd_dendro[[j]]), 3)
  
  dendextend::tanglegram(dend1 = ctd_dendro[[i]], 
                         dend2 = ctd_dendro[[j]],
                         main_left = i,
                         main_right =  j,
                         main = paste("Entanglement =", tangleidx ),
                         sub = paste("Entanglement =", tangleidx ),
                         #cex_main = 100,
                         #cex_sub = 100,
                         common_subtrees_color_branches = TRUE
                         )
  
}

tangleidx <- foreach(i = dendro_comb[, 1], 
                      j = dendro_comb[, 2], 
                      .final = function(l) setNames(l, dendro_names)) %do% {
  
  round(dendextend::entanglement(ctd_dendro[[i]], ctd_dendro[[j]]), 3)

}

#' Run directly in the console
for (i in names(ctd_tangle)) {
  pdf(path(proj$img, paste("tangle", i, sep = "-"), ext = "pdf"), 
      width = 40, 
      height = 20)
  plot(ctd_tangle[[i]])
  dev.off()
}
```





* Correlation cophenetic ~ dist
Compute some dendogram and compare them

```{r}
# Compute cophentic distance 
ctd_coph <- vector(mode = "list")
for (i in names(ctd_hclust)){
  ctd_coph[[i]] <- cophenetic(ctd_hclust[[i]])   
}

ctd_coph_cor <- vector(mode = "list")
for (i in names(ctd_hclust)){
  ctd_coph_cor[[i]] <- cor(ctd_dist, ctd_coph[[i]])   
}

ctd_coph_cor
```

* Dendogram (horizontal) fviz_dend

```{r}
pic <- factoextra::fviz_dend(ctd_hclust$ward.D2, 
                      horiz = TRUE, 
                      cex = 0.75)

ggsave(filename = path(proj$img, "clara_analyais_gap_statistic", ext = "pdf"), 
       plot = pic, 
       width=15, height=40)
#' Run directly in the console
# pdf("dendrogram.pdf", width=15, height=30) 
# print(p) 
# dev.off()

```

* Choosing among groups and algorithms

```{r}
smu1_stu <- vector(mode = "numeric")
smu1_stu[["stu"]] <- RPostgres::dbGetQuery(conGIS, glue_sql(
  "WITH smu1_stu AS (SELECT DISTINCT stu_id
    FROM smu1_stu
    WHERE stupc iS NOT NULL
    ORDER BY stu_id
    )
  SELECT count(*)
  FROM smu1_stu
  ", 
  .con = conGIS)
  ) %>% 
  flatten_dbl()
```

```{r}
ctd_valid <- clValid::clValid(ctd_mat, 
                 nClust = seq(from = 35, to = 155, by = 30), 
                 clMethods = c("hierarchical", "agnes"),
                 method = c("ward"),
                 metric = "euclidean",
                 validation = c("internal", "stability") 
                 )

summary(ctd_valid)
```


Validation:

* pvclust

```{r}
#' Create FORK cluster limited to 6 cores
registerDoSEQ()

ctd_pv <- pvclust(parallel = cores, t(ctd_mat), method.hclust = "ward.D2",
    method.dist = "correlation", nboot = 500, iseed = 5432)

# Reset to standard cluster
parallel::stopCluster(cores)
registerDoParallel(cores = 6L)

```




```{r}
pic <- factoextra::fviz_nbclust(smu$sub, cluster::clara, k.max = max(search_space), method = "gap_stat", nboot = 10) +
  labs(subtitle = "Gap statistic method")
ggsave(filename = path(proj$img, "gap_statistic", ext = "svg"), plot = pic)

NbClustPar
```


```{r}
grp <- cutree(res.hc, k = 4)
fviz_dend(res.hc, k = 4, # Cut in four groups
cex = 0.5, # label size 
k_colors = c("#2E9FDF", "#00AFBB", "#E7B800", "#FC4E07"), 
color_labels_by_k = TRUE, # color labels by groups 
rect = TRUE # Add rectangle around groups 
)

fviz_cluster(list(data = df, cluster = grp),
palette = c("#2E9FDF", "#00AFBB", "#E7B800", "#FC4E07"), 
ellipse.type = "convex", # Concentration ellipse 
repel = TRUE, # Avoid label overplotting (slow) 
show.clust.cent = FALSE, 
ggtheme = theme_minimal())
```
 

Validation:

* pvclust
```{r}
# Utilizar fork
parPvclust(cl=NULL, data, method.hclust = "average",

method.dist = "correlation", nboot = 1000, iseed = NULL)
```




## Comparativa 

## Clasificación definitiva


