---
title: "GIS bridges"
author: "Alberto Lázaro-López"
date: "22th March 2019"
output: 
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
abstract: Filter of a DEM to be used in Digital Soil Mapping. 
---

```{r core, include=FALSE }
for (.i in paste0("src/", c("core", "raster"), ".R" ) ) source(.i)
```


# SAGA

## RSAGA
Install and load the [RSAGA](https://github.com/r-spatial/RSAGA) package from CRAN.

```{r rsaga}
#install.packages("RSAGA")
```

In order to use SAGA, we need to set up an environment variable which will be pass to each geoprocess calling.
We can either specify the cmd and modules paths or supply the main folder of installation. 
I choose the latter, as it's safe againts version uppgrades.

```{r rsaga_env, echo=FALSE}
library(RSAGA)

env <- rsaga.env(root = "/usr/local/opt/saga-gis",
                 parallel = TRUE)
```

The saga_cmd, the core behind of RSAGA, just read data directly from file: SAGA specific file format for rasters and Shapefile for vectors.
So, for every process, one needs to follow a read / write workflow in the workspace folder.

When we are working with rasters from a PostGIS database, there is an option to avoid the R step and to directly convert them to SAGA through GDAL.
RSAGA package has a function for that:

```{r rsaga_write}
pgis <- pgiscon("mdt", "dsm", .db = "tezisdb")

rsaga.import.gdal(in.grid = pgis,
                  out.grid = "test",
                  env = env)
```

It does it in two septs though, and its notes suggests that a direct translation using GDAL is a faster approach.
That's why we need to call gdal_translate from system

```{r gdal_write}
glue('gdal_translate -of SAGA -a_srs EPSG:25830 "{pgis}" {file.path(getwd(), "_tmp/dem.sdat")}'  ) %>% system()
```

Testing a geoprocess 

```{r geoprocess_test}
rsaga.geoprocessor(lib = "ta_morphometry", module = "Slope, Aspect, Curvature",
                   param = list(ELEVATION = file.path(getwd(), "_tmp/dem.sgrd"), 
                                SLOPE = file.path(getwd(), "_tmp/slope.sgrd"),
                                UNIT_SLOPE = "1"),
                   env = env)
```




## Rsagacmd

Install the [rsagacmd](https://github.com/stevenpawley/Rsagacmd) package.

```{r rsagacmd}
#devtools::install_github("stevenpawley/Rsagacmd")
```

Automatic build of all SAGA functions
```{r rsagacmd_func}
library(Rsagacmd)

saga <- saga_gis(saga_bin = "/usr/local/opt/osgeo-saga/bin/saga_cmd")
```

`Rsagacmd` has a more R-notion approach, as all functions for SAGA area readibly within R in a form of nested lists, it suggests parameters of functions and conversion between filetypes is made in the background transparently to the user. 
Whereas `RSAGA`seems to use a lowerlevel language because you need to take care of filetypes and to implement correctly every function.

Testing a geoprocess

```{r rsagacmd_proc}
# Getting DEM for test
pgis <- pgiscon("mdt", "_dsm", .db = "tezisdb", .where = glue::glue('smu1_id~ $${smu1}$$ ', smu1='J04') )

dem <- readGDAL(pgis, silent = TRUE)
  brick() %>% 
  raster(layer="band6")

# Geoprocessing
tri <- saga$ta_morphometry$Terrain_Ruggedness_Index_TRI(DEM = dem)
```

It took a long time to get processed a small piece of DEM. It seems `RSAGA` is more efficient and let you pick desired outputs, so it's going to be the standard package in a normal basis. 




# GRASS

## RGRASS7

Install the `rgrass7` package.

```{r rgrass7_ins}
#install.packages("rgrass7")
```

GRASS GIS demand to work under a GRASS database, which collects locations with at least a mapset.
During standard workflow, a temporal database can be set up. 
Whereas during development on an fix area, a location with a mapset can be buildreuse in order to avoid the configuration process.

A minimum requirement is to configure the Coordinate Reference System of the location before loading any data.

```{r grass_setup}
library(rgrass7)

grass_path <- initGRASS(gisBase = "/usr/local/opt/osgeo-grass/grass-base", 
          home = tempdir(),
          gisDbase = tempdir(), 
          location = "rstudio", 
          mapset = "PERMANENT", 
          override = TRUE )

execGRASS("g.proj", flags = c("c", "quiet"),
          epsg = 3042 )

gmeta()
```

After that any data source accepted by GDAL can be used, which is quite powerful!. 
This is a way to transfer data from a PostGIS database to GRASS database using a GDAL string.
The result can be checked with the corresponding GRASS module. 

When the first raster data is ready, the [Computational region][^reg] can be set.
This is an important concept, as every calculation on raster will be reduced to this area. 

[^reg]: https://grasswiki.osgeo.org/wiki/Unleash_the_power_of_GRASS_GIS_at_US-IALE_2017#Scripting_with_R

```{r pgis_grass}
pgis <- pgiscon("mdt", "dsm", .db = "tezisdb")

execGRASS("r.in.gdal", flags = c("o"),
          input = pgis,
          output = "dem"  )

execGRASS("g.list", flags = c(),
          type="all")

execGRASS("g.region", raster="dem", flags=c("a","p"))
```

When data isn't in the CRS of the project, `r.import` module can be used, which reprojects on the fly.

```{r reproj}
execGRASS("r.import", flags = c("o"),
          input = pgis,
          output = "dem2"  )

execGRASS("g.list", flags = c(),
          type="all")

execGRASS("g.remove", flags = c("f"),
          type = "raster",
          name = "dem2"  )
```

Generally, executing a GRASS geoprocess is done with `execGRASS` which produces the message as a system call.

```{r grass_proc}
execGRASS("r.slope.aspect", flags = c("n"),
          elevation="dem",
          slope="slope",
          aspect="aspect",
          format="degrees",
          min_slope=5.0)
```

`doGRASS` is another alternative to `execGRASS`, which executes a process but gives a string and, therefore, it can be used within a loop.

After the process, the new raster bands can be read into R session and plot to check.

```{r}
geograss_raw <- rgrass7::readRAST(c("slope", "aspect"))
geograss <- brick(geograss_raw)

plot(geograss$slope)
```

Or can be exported to an intermediate file to be loaded into a PostGIS database

```{r}
output <- glue('_tmp/{name}.tiff', name="slope" )

execGRASS("r.out.gdal", flags = c(),
          input="slope",
          output=output,
          format="GTiff")

execGRASS("r.out.gdal", flags = c("l"))
```


When we want to pass data ready within the R session, it can be loaded into GRASS database as:

```{r}
dem_grid <- as(dem, 'SpatialGridDataFrame')
writeRAST(dem_grid, "dem_j04c")

execGRASS("g.list", flags = c(),
          type="all")
```

Vector data too

```{r}
cons <- dbConnect(RPostgres::Postgres(),  
                 dbname="tezisdb", host="localhost", port=5432, 
                 user={keyring::key_list("psql-su")[1,2]}, 
                 password=keyring::key_get("psql-su"))
space <- st_read(cons,
  query = "select * from do_cbj")

b_box = st_bbox(space) 
execGRASS("g.region", flags = c("quiet"), 
          n = as.character(b_box["ymax"]), s = as.character(b_box["ymin"]), 
          e = as.character(b_box["xmax"]), w = as.character(b_box["xmin"]), 
          res = "1")
```








# QGIS

## RQGIS3

Install and load the `RQGIS3` package.
```{r rqgis3, echo=FALSE}
devtools::install_github("jannes-m/RQGIS3", ref = "mac" )
```


```{r rqgis3_env, echo=FALSE}
library(RQGIS3)

set_env(root = "/usr/local/Cellar/qgis/3.6.0_4/QGIS.app", dev = FALSE, new = TRUE)
open_app()
```

When running the last function RStudio crashes. 
If this is done in a Terminal session, it works with an adjustment: it will ask you to duplicate a symlink of actual Python 3 as version 3.6.

```{r}
alg <- find_algorithms(search_term = "slope", name_only = TRUE)
get_usage(alg = alg[2])
param <- get_args_man(alg = alg[2])
get_options(alg[2])

param$elevation <- dem
```






